---
title: "Data Analytics & Results"
subtitle: "Evidence-Based Insights into IE Programs"
sidebar: false
---

### Key Takeaways
* **Comprehensive Data:** 2018-2025 multi-year data extracted via `thestats`.
* **Analytical Approach:** Moving from descriptive statistics to predictive insights.
* **Faculty Factor:** Measuring the real impact of academic staff size on student rankings.

---

<div class="glass-card">
## Exploratory Analysis
*(Burada harika ggplot2 grafiklerimiz ve yorumlarımız olacak)*
</div>

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
# Analiz kodları buraya gelecek
### Key Takeaways

* **Comprehensive Data:** 2018-2025 multi-year data extracted via `thestats`.
* **Analytical Approach:** Moving from descriptive statistics to predictive insights.
* **Faculty Factor:** Measuring the real impact of academic staff size on student rankings.
* **Regional Dynamics:** Identifying how geographical location influences program popularity.

---

<div class="glass-card">
## 1. Introduction to Analysis
This section presents our exploratory data analysis and findings. By leveraging the power of `tidyverse` and `ggplot2`, we uncover the trends that define Industrial Engineering education in Turkey.
</div>

## 2. Preliminary Data Exploration
*(Buraya veri çekme ve ilk görselleştirme kodlarımızı ekleyeceğiz.)*

```{r}
#| label: setup
#| message: false
#| warning: false
#| code-fold: true

library(tidyverse)
# library(thestats) # Veri çekme aşamasında aktif edilecek