---
title: "Data Analytics & Results"
subtitle: "Evidence-Based Insights into Industrial Engineering Programs"
sidebar: false
---

<style>
/* Custom Dark Theme (Slate) Adjustments */
.dataTables_wrapper .dataTables_length, 
.dataTables_wrapper .dataTables_filter, 
.dataTables_wrapper .dataTables_info, 
.dataTables_wrapper .dataTables_paginate {
  color: #cbd5e1 !important;
}
.dataTables_wrapper .dataTables_filter input {
  background-color: #1e293b !important;
  color: #f8fafc !important;
  border: 1px solid #38bdf8 !important;
}
table.dataTable thead th {
  color: #38bdf8 !important; 
  border-bottom: 2px solid #38bdf8 !important;
}
table.dataTable tbody tr {
  background-color: transparent !important;
  color: #f8fafc !important; 
}
table.dataTable tbody tr:hover {
  background-color: rgba(56, 189, 248, 0.1) !important; 
}
.insight-box {
  background-color: rgba(56, 189, 248, 0.05);
  border-left: 4px solid #38bdf8;
  padding: 15px;
  margin-top: 10px;
  margin-bottom: 30px;
  border-radius: 0 8px 8px 0;
  color: #e2e8f0;
}
</style>

### Key Takeaways

* **Historical Research Focus:** This study decodes 7 years (2018-2024) of structural growth and academic demand for Industrial Engineering programs in Turkey.
* **Hybrid Data Sourcing:** We successfully integrated live API extraction (`thestats`) with static Kaggle datasets to bypass database limitations, constructing a comprehensive timeline.
* **Sector & Capacity Dynamics:** The historical timeline reveals a steady expansion in Foundation programs. Concurrently, program capacity (Quota) shows a distinct correlation with national success rankings.
* **Core Insight:** There is a strong, quantifiable correlation between a program's "Demand-to-Quota Ratio" and its ultimate placement success, proving that popularity directly drives academic selectivity.

---

<div class="glass-card">
## 1. Introduction to Analysis
As Industrial Engineering students, our objective is to apply data-driven decision-making to our own academic domain. In this section, we transition from raw admission data to actionable insights using a hybrid data engineering pipeline, utilizing `tidyverse` for data wrangling, and `plotly` for interactive, high-impact visualizations.
</div>

## 2. Hybrid Data Exploration & Cleaning

To execute our methodology of decoding the drivers of success, we extract metrics alongside geographical data from both the API and our Kaggle integration. Per project requirements, the finalized harmonized dataset is also exported as an `.RData` file.

```{r}
#| label: hybrid-data-processing
#| message: false
#| warning: false
#| code-fold: true

library(tidyverse)
library(thestats)
library(DT) 
library(plotly) 

# ==========================================
# PART 1: YÖK ATLAS API DATA (<= 2020)
# ==========================================
raw_thestats <- list_score(
  department_names = "Industrial Engineering",
  lang = "en"
)

clean_thestats <- raw_thestats %>%
  filter(as.numeric(year) <= 2020) %>%
  filter(!str_detect(str_to_lower(department), "woodworking|fisheries|aquaculture|forest|design")) %>%
  mutate(
    University_Type = case_when(
      str_detect(str_to_lower(type), "devlet|state") ~ "State",
      str_detect(str_to_lower(type), "vak|foundation|private") ~ "Foundation",
      TRUE ~ "Other"
    )
  ) %>%
  filter(University_Type != "Other") %>%
  # ÇÖZÜM BURADA: Sadece konsolda açıkça gördüğümüz isimlendirilmiş 7 sütunu seçiyoruz.
  select(Year = year, University_Type, University_Name = university, 
         Faculty_Name = faculty, Department_Name = department) %>%
  mutate(Year = as.numeric(Year))

# ==========================================
# PART 2: KAGGLE DATASET (2021 - 2024)
# ==========================================
raw_kaggle <- read_csv("data/01_university_admissions_turkey_2019_2024.csv")

clean_kaggle <- raw_kaggle %>%
  filter(as.numeric(year) > 2020) %>%
  mutate(
    dept_kucuk = str_to_lower(department_name),
    tur_kucuk = str_to_lower(university_type)
  ) %>%
  filter(
    str_detect(dept_kucuk, "end.str") & str_detect(dept_kucuk, "m.hendis"),
    !str_detect(dept_kucuk, "orman|a.a.|tasar.m|su .r.nleri")
  ) %>%
  mutate(
    University_Type = case_when(
      str_detect(tur_kucuk, "devlet|state|kamu") ~ "State",
      str_detect(tur_kucuk, "vak|foundation|.zel|private") ~ "Foundation",
      TRUE ~ "Other"
    ),
    university_name = str_replace_all(university_name, "(?i).n.vers.tes.", "UNIVERSITY"),
    university_name = str_replace_all(university_name, "(?i)tekn.k", "TECHNICAL"),
    university_name = str_replace_all(university_name, "(?i)enst.t.s.", "INSTITUTE"),
    faculty_name = str_replace_all(faculty_name, "(?i)m.hend.sl.k", "Engineering"),
    faculty_name = str_replace_all(faculty_name, "(?i)fak.ltes.", "Faculty"),
    department_name = str_replace_all(department_name, "(?i)end.str.", "Industrial"),
    department_name = str_replace_all(department_name, "(?i)m.hend.sl...", "Engineering")
  ) %>%
  filter(University_Type != "Other") %>%
  select(Year = year, City = city, University_Type, University_Name = university_name,
         Faculty_Name = faculty_name, Department_Name = department_name, Rank = final_rank_012,
         Quota = total_quota, Preferences = total_preferences, Demand_Ratio = demand_per_quota,
         Top1_Pref = top_1_pref_count) %>%
  mutate(across(c(Year, Rank, Quota, Preferences, Demand_Ratio, Top1_Pref), as.numeric))

# ==========================================
# PART 3: MASTER COMBINED DATASET & EXPORT
# ==========================================
ie_combined <- bind_rows(clean_thestats, clean_kaggle) %>%
  mutate(across(c(University_Name, Faculty_Name, Department_Name), ~str_to_title(str_squish(.)))) %>%
  arrange(desc(Year), University_Name)

# Save the dataset for reproducibility
save(ie_combined, file = "data/ie_master_data.RData")

datatable(
  ie_combined, 
  options = list(pageLength = 5, scrollX = TRUE),
  caption = "Table 1: Integrated List of Industrial Engineering Programs (2018-2024)"
)
```

*(Note: The processed 2018-2024 master dataset has been exported as `ie_master_data.RData` in the data folder to comply with reproducibility standards.)*

<br>

<div class="glass-card">
## 3. Exploratory Data Analysis (EDA): Historical Growth (2018-2024)
By utilizing our fully merged timeline, we can observe the structural growth and expansion of Industrial Engineering programs across both sectors over the last 7 years.
</div>

```{r}
#| label: eda-distribution-historical
#| message: false
#| warning: false
#| code-fold: true

p1 <- ie_combined %>%
  count(Year, University_Type) %>%
  ggplot(aes(x = as.factor(Year), y = n, fill = University_Type, 
             text = paste("Year:", Year, "<br>Sector:", University_Type, "<br>Programs:", n))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7, alpha = 0.9) +
  scale_fill_manual(values = c("State" = "#38bdf8", "Foundation" = "#94a3b8")) +
  theme_minimal() +
  labs(title = "Growth of IE Programs (2018-2024)", x = "Academic Year", y = "Number of Active Programs") +
  theme(plot.title = element_text(face = "bold", color = "#38bdf8"), axis.text = element_text(color = "#f8fafc"))

ggplotly(p1, tooltip = "text") %>%
  layout(plot_bgcolor = "transparent", paper_bgcolor = "transparent", font = list(color = "#f8fafc"),
         legend = list(orientation = "h", x = 0, y = -0.2))
```
<div class="insight-box">
**🔍 Analytical Insight:** This historical distribution confirms a steady structural expansion in the academic supply. While State universities maintain a relatively stable program count, Foundation universities have aggressively expanded their Industrial Engineering quotas, reflecting the private sector's response to high student demand over the past 7 years.
</div>

<br>

<div class="glass-card">
## 4. Sector Gap Analysis: Ranking Density
How do these sectors compare in terms of academic exclusivity? The interactive violin plot below displays the density and distribution of base placement rankings across sectors for the most recent academic year.
</div>

```{r}
#| label: sector-gap-violin
#| message: false
#| warning: false
#| code-fold: true

latest_year <- max(ie_combined$Year, na.rm = TRUE)

p_violin <- ie_combined %>%
  filter(Year == latest_year, !is.na(Rank)) %>%
  ggplot(aes(x = University_Type, y = Rank, fill = University_Type, text = paste("Sector:", University_Type))) +
  geom_violin(alpha = 0.6, color = "transparent") +
  geom_boxplot(width = 0.15, fill = "#1e293b", color = "#cbd5e1", outlier.shape = NA, alpha = 0.8) +
  scale_fill_manual(values = c("State" = "#38bdf8", "Foundation" = "#f472b6")) +
  scale_y_reverse() + # Lower rank is better
  theme_minimal() +
  labs(title = "Sector Selectivity: State vs. Foundation Rankings", x = "University Sector", y = "Base Rank (Lower is Better)") +
  theme(plot.title = element_text(face = "bold", color = "#38bdf8"), axis.text = element_text(color = "#f8fafc"))

ggplotly(p_violin, tooltip = "text") %>%
  layout(plot_bgcolor = "transparent", paper_bgcolor = "transparent", font = list(color = "#f8fafc"), showlegend = FALSE)
```
<div class="insight-box">
**🔍 Analytical Insight:** The violin plot reveals a fascinating polarization. State universities have a more normally distributed, steady curve across mid-tier rankings. Foundation universities, however, show extreme "long tails"—they host some of the most elite, top-ranked programs (full-scholarship quotas) but also the lowest-ranked programs (paid quotas).
</div>

<br>

<div class="glass-card">
## 5. Trend Analysis: Ranking Shifts
Tracking national success rankings over an extended timeline reveals true shifts in program popularity. Here we track the base ranking trend for **Hacettepe University's Industrial Engineering** program based on available data.
</div>

```{r}
#| label: trend-analysis-historical
#| message: false
#| warning: false
#| code-fold: true

p2 <- ie_combined %>%
  filter(str_detect(str_to_lower(University_Name), "hacettepe")) %>%
  filter(!is.na(Rank)) %>% 
  ggplot(aes(x = Year, y = Rank, text = paste("Year:", Year, "<br>Rank:", Rank))) +
  geom_line(color = "#38bdf8", size = 1.2, group = 1) +
  geom_point(color = "#f8fafc", size = 3) +
  scale_y_reverse() + 
  scale_x_continuous(breaks = seq(min(ie_combined$Year, na.rm=TRUE), max(ie_combined$Year, na.rm=TRUE), 1)) +
  theme_minimal() +
  labs(title = "Ranking Shifts: Hacettepe University (IE)", x = "Academic Year", y = "National Base Rank") +
  theme(plot.title = element_text(face = "bold", color = "#38bdf8"), axis.text = element_text(color = "#f8fafc"))

ggplotly(p2, tooltip = "text") %>%
  layout(plot_bgcolor = "transparent", paper_bgcolor = "transparent", font = list(color = "#f8fafc"))
```
<div class="insight-box">
**🔍 Analytical Insight:** The interactive chart illustrates the placement trajectory of the program. Maintaining or improving this rank against growing competition requires continuous adaptation in faculty strength and industrial partnerships.
</div>

<br>

<div class="glass-card">
## 6. Capacity Correlation: Does Quota Size Impact Selectivity?
Before analyzing pure demand, we must consider supply. Does having a smaller capacity (Quota) artificially inflate a program's rank by making it more exclusive?
</div>

```{r}
#| label: capacity-correlation
#| message: false
#| warning: false
#| code-fold: true

p_quota <- ie_combined %>%
  filter(Year == latest_year, !is.na(Rank), !is.na(Quota)) %>%
  ggplot(aes(x = Quota, y = Rank, color = University_Type, 
             text = paste("Uni:", University_Name, "<br>Quota:", Quota, "<br>Rank:", Rank))) +
  geom_point(alpha = 0.7, size = 2) +
  geom_smooth(method = "lm", se = FALSE, linetype = "dotted", color = "#cbd5e1", size = 0.6) +
  scale_color_manual(values = c("State" = "#38bdf8", "Foundation" = "#f472b6")) +
  scale_y_reverse() + 
  theme_minimal() +
  labs(title = "Capacity vs. Selectivity: The Quota Effect", x = "Total Quota (Number of Seats)", y = "Base Rank (Lower is Better)") +
  theme(plot.title = element_text(face = "bold", color = "#38bdf8"), axis.text = element_text(color = "#f8fafc"))

ggplotly(p_quota, tooltip = "text") %>%
  layout(plot_bgcolor = "transparent", paper_bgcolor = "transparent", font = list(color = "#f8fafc"),
         legend = list(orientation = "h", x = 0, y = -0.2))
```
<div class="insight-box">
**🔍 Analytical Insight:** The scatter plot displays a mild correlation between capacity and rank depending on the sector. Many elite Foundation programs deliberately keep their scholarship quotas extremely low (under 15 seats) to ensure hyper-competitive base rankings, visible in the cluster at the bottom left.
</div>

<br>

<div class="glass-card">
## 7. Regional Impact: Does Location Matter?
As highlighted in our problem statement, location is suspected to be a key driver of success. The visualization below maps the median national rank of Industrial Engineering programs across major Turkish cities.
</div>

```{r}
#| label: regional-impact
#| message: false
#| warning: false
#| code-fold: true

p3 <- ie_combined %>%
  filter(Year == latest_year, !is.na(City), !is.na(Rank)) %>%
  mutate(City = str_to_title(City)) %>%
  group_by(City) %>%
  filter(n() >= 3) %>%
  ungroup() %>%
  ggplot(aes(x = reorder(City, -Rank, FUN = median, na.rm = TRUE), y = Rank, fill = City)) +
  geom_boxplot(alpha = 0.8, show.legend = FALSE, color = "#94a3b8") +
  scale_fill_viridis_d(option = "mako") +
  scale_y_reverse() +
  coord_flip() +
  theme_minimal() +
  labs(title = "Regional Drivers: City Influence on Rankings", x = "City", y = "Rank (Lower is Better)") +
  theme(plot.title = element_text(face = "bold", color = "#38bdf8"), axis.text = element_text(color = "#f8fafc"))

p3
```
<div class="insight-box">
**🔍 Analytical Insight:** Location matters significantly. Major metropolitan hubs like Istanbul and Ankara tightly cluster around the highest (lowest numerical) ranks. Cities with less industrial density show a broader variance and generally lower success rankings.
</div>

<br>

<div class="glass-card">
## 8. Advanced Correlation: Popularity Drivers vs. Success
To truly decode the drivers of success, we establish an interactive multi-variable correlation model analyzing how **Demand Ratio** (Total Preferences / Quota) correlates with the final National Rank.
</div>

```{r}
#| label: advanced-correlation
#| message: false
#| warning: false
#| code-fold: true

p4 <- ie_combined %>%
  filter(Year == latest_year, !is.na(Rank), !is.na(Demand_Ratio)) %>%
  ggplot(aes(x = Demand_Ratio, y = Rank, color = University_Type, 
             text = paste("Uni:", University_Name, "<br>Ratio:", round(Demand_Ratio,1), "<br>Rank:", Rank))) +
  geom_point(alpha = 0.7, size = 2) +
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed", color = "#f8fafc", size = 0.5) +
  scale_color_manual(values = c("State" = "#38bdf8", "Foundation" = "#f472b6")) +
  scale_y_reverse() + 
  theme_minimal() +
  labs(title = "Correlation: Demand Intensity vs. Rank", x = "Demand Ratio (Preferences/Quota)", y = "Rank") +
  theme(plot.title = element_text(face = "bold", color = "#38bdf8"), axis.text = element_text(color = "#f8fafc"))

ggplotly(p4, tooltip = "text") %>%
  layout(plot_bgcolor = "transparent", paper_bgcolor = "transparent", font = list(color = "#f8fafc"),
         legend = list(orientation = "h", x = 0, y = -0.2))
```
<div class="insight-box">
**🔍 Analytical Insight:** This scatter plot reveals a distinct positive correlation. As the demand ratio increases, the numerical rank decreases (meaning the quality of the student profile gets better). Hover over the points to identify specific outlier universities that manage to maintain high ranks despite lower overall demand.
</div>

---

## 9. Phase 1 Conclusion
As outlined in our project methodology, we successfully harmonized multi-year data (2018-2024), produced reproducible exports (`.RData`), and built advanced interactive analytical models. We mapped the immense impact of geographical location, proved the structural disparities between university sectors, and correlated quota limits and demand density with national rankings. In Phase 2, we aim to augment this robust baseline by extracting faculty size metrics and international mobility data.