---
title: "Data Analytics & Results"
subtitle: "Evidence-Based Insights into Industrial Engineering Programs"
sidebar: false
---

<style>
/* DT Tablosunu Karanlık Temaya (Slate) Uyduran CSS */
.dataTables_wrapper .dataTables_length, 
.dataTables_wrapper .dataTables_filter, 
.dataTables_wrapper .dataTables_info, 
.dataTables_wrapper .dataTables_paginate {
  color: #cbd5e1 !important;
}
.dataTables_wrapper .dataTables_filter input {
  background-color: #1e293b !important;
  color: #f8fafc !important;
  border: 1px solid #38bdf8 !important;
}
table.dataTable thead th {
  color: #38bdf8 !important; 
  border-bottom: 2px solid #38bdf8 !important;
}
table.dataTable tbody tr {
  background-color: transparent !important;
  color: #f8fafc !important; 
}
table.dataTable tbody tr:hover {
  background-color: rgba(56, 189, 248, 0.1) !important; 
}
</style>

### Key Takeaways

* **Hybrid Data Sourcing:** Successfully integrated live API extraction (`thestats`) with static datasets (Kaggle) to construct a comprehensive timeline.
* **National Trends:** Clear shifts identified in placement capacities and base rankings across Turkish universities.
* **Regional & Demand Correlations:** Advanced modeling reveals strong relationships between a program's geographical location, its demand-to-quota ratio, and its ultimate national success ranking.

---

<div class="glass-card">
## 1. Introduction to Analysis
As Industrial Engineering students, our objective is to apply data-driven decision-making to our own academic domain. In this section, we transition from raw admission data to actionable insights using a hybrid data engineering pipeline, utilizing `tidyverse` for data wrangling and `ggplot2` for high-impact visualizations.
</div>

## 2. Hybrid Data Exploration & Cleaning

To execute our methodology of decoding the drivers of success, we extract deep numeric metrics (Rankings, Quotas, Preferences, Demand Ratios, and Top-Choice counts) alongside geographical data from our Kaggle integration.

```{r}
#| label: hybrid-data-processing
#| message: false
#| warning: false
#| code-fold: true

library(tidyverse)
library(thestats)
library(DT) 

# ==========================================
# PART 1: YÖK ATLAS API DATA (<= 2020)
# ==========================================
raw_thestats <- list_score(
  department_names = "Industrial Engineering",
  lang = "en"
)

clean_thestats <- raw_thestats %>%
  filter(as.numeric(year) <= 2020) %>%
  filter(!str_detect(str_to_lower(department), "woodworking|fisheries|aquaculture|forest|design")) %>%
  mutate(
    University_Type = case_when(
      str_detect(str_to_lower(type), "devlet|state") ~ "State",
      str_detect(str_to_lower(type), "vak|foundation|private") ~ "Foundation",
      TRUE ~ "Other"
    )
  ) %>%
  filter(University_Type != "Other") %>%
  select(
    Year = year,
    University_Type,
    University_Name = university,
    Faculty_Name = faculty,
    Department_Name = department
  ) %>%
  mutate(Year = as.numeric(Year))

# ==========================================
# PART 2: KAGGLE DATASET (2021 - 2024) - KAPSAMLI VERİ ÇEKİMİ
# ==========================================
raw_kaggle <- read_csv("data/01_university_admissions_turkey_2019_2024.csv")

clean_kaggle <- raw_kaggle %>%
  filter(as.numeric(year) > 2020) %>%
  mutate(
    dept_kucuk = str_to_lower(department_name),
    tur_kucuk = str_to_lower(university_type)
  ) %>%
  filter(
    str_detect(dept_kucuk, "end.str") & str_detect(dept_kucuk, "m.hendis"),
    !str_detect(dept_kucuk, "orman|a.a.|tasar.m|su .r.nleri")
  ) %>%
  mutate(
    University_Type = case_when(
      str_detect(tur_kucuk, "devlet|state|kamu") ~ "State",
      str_detect(tur_kucuk, "vak|foundation|.zel|private") ~ "Foundation",
      TRUE ~ "Other"
    ),
    university_name = str_replace_all(university_name, "(?i).n.vers.tes.", "UNIVERSITY"),
    university_name = str_replace_all(university_name, "(?i)tekn.k", "TECHNICAL"),
    university_name = str_replace_all(university_name, "(?i)enst.t.s.", "INSTITUTE"),
    faculty_name = str_replace_all(faculty_name, "(?i)m.hend.sl.k", "Engineering"),
    faculty_name = str_replace_all(faculty_name, "(?i)fak.ltes.", "Faculty"),
    department_name = str_replace_all(department_name, "(?i)end.str.", "Industrial"),
    department_name = str_replace_all(department_name, "(?i)m.hend.sl...", "Engineering")
  ) %>%
  filter(University_Type != "Other") %>%
  # HEDEFLENEN YENİ VERİLER: Şehir, Talep Oranı, İlk Tercih Edenler
  select(
    Year = year,
    City = city,
    University_Type,
    University_Name = university_name,
    Faculty_Name = faculty_name,
    Department_Name = department_name,
    Rank = final_rank_012,
    Quota = total_quota,
    Preferences = total_preferences,
    Demand_Ratio = demand_per_quota,
    Top1_Pref = top_1_pref_count
  ) %>%
  mutate(
    Year = as.numeric(Year),
    Rank = as.numeric(Rank),
    Quota = as.numeric(Quota),
    Preferences = as.numeric(Preferences),
    Demand_Ratio = as.numeric(Demand_Ratio),
    Top1_Pref = as.numeric(Top1_Pref)
  )

# ==========================================
# PART 3: MASTER COMBINED DATASET
# ==========================================
ie_combined <- bind_rows(clean_thestats, clean_kaggle) %>%
  mutate(
    University_Name = str_to_title(str_squish(University_Name)),
    Faculty_Name = str_to_title(str_squish(Faculty_Name)),
    Department_Name = str_to_title(str_squish(Department_Name))
  ) %>%
  arrange(desc(Year), University_Name)

datatable(
  ie_combined, 
  options = list(pageLength = 5, scrollX = TRUE),
  caption = "Table 1: Integrated List of Industrial Engineering Programs (2018-2024)"
)
```

<br>

<div class="glass-card">
## 3. Exploratory Data Analysis (EDA): Structural Distribution
With our seamlessly integrated master dataset, we visualize the structural distribution of Industrial Engineering programs for the most recent academic year.
</div>

```{r}
#| label: eda-distribution
#| message: false
#| warning: false
#| code-fold: true

latest_year <- max(ie_combined$Year, na.rm = TRUE)

ie_combined %>%
  filter(Year == latest_year) %>%
  count(University_Type) %>%
  ggplot(aes(x = University_Type, y = n, fill = University_Type)) +
  geom_bar(stat = "identity", width = 0.5, show.legend = FALSE, alpha = 0.9) +
  scale_fill_manual(values = c("State" = "#38bdf8", "Foundation" = "#94a3b8")) +
  geom_text(aes(label = n), vjust = -0.5, fontface = "bold", color = "white", size = 5) +
  theme_minimal() +
  labs(
    title = paste("Distribution of IE Programs in", latest_year),
    subtitle = "Active Programs in State vs. Foundation Universities",
    x = "University Sector",
    y = "Number of Active Programs"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 16, color = "#38bdf8"),
    plot.subtitle = element_text(size = 12, color = "#cbd5e1", margin = margin(b = 15)),
    axis.text = element_text(color = "#f8fafc", size = 11),
    axis.title = element_text(face = "bold", color = "#f8fafc", size = 12),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(color = "gray", linetype = "dotted")
  )
```

<br>

<div class="glass-card">
## 4. Trend Analysis: Ranking Shifts in Top Universities
Tracking national success rankings over time reveals shifts in program popularity. Here we track the base ranking trend for Hacettepe University's Industrial Engineering program.
</div>

```{r}
#| label: trend-analysis
#| message: false
#| warning: false
#| code-fold: true

ie_combined %>%
  filter(str_detect(str_to_lower(University_Name), "hacettepe")) %>%
  filter(!is.na(Rank)) %>% 
  ggplot(aes(x = Year, y = Rank)) +
  geom_line(color = "#38bdf8", size = 1.5) +
  geom_point(color = "#f8fafc", size = 4) +
  geom_text(aes(label = Rank), vjust = -1.5, color = "#cbd5e1", fontface = "bold") +
  scale_y_reverse() + 
  scale_x_continuous(breaks = seq(2021, 2024, 1)) +
  theme_minimal() +
  labs(
    title = "Ranking Shifts: Hacettepe University (IE)",
    subtitle = "Base Placement Rank Trend (Lower is Better)",
    x = "Academic Year",
    y = "National Base Rank"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 16, color = "#38bdf8"),
    plot.subtitle = element_text(size = 12, color = "#cbd5e1", margin = margin(b = 15)),
    axis.text = element_text(color = "#f8fafc", size = 11),
    axis.title = element_text(face = "bold", color = "#f8fafc", size = 12),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(color = "gray", linetype = "dotted")
  )
```

<br>

<div class="glass-card">
## 5. Regional Impact: Does Location Matter?
As highlighted in our problem statement, location is suspected to be a key driver of success. The visualization below maps the median national rank of Industrial Engineering programs across major Turkish cities.
</div>

```{r}
#| label: regional-impact
#| message: false
#| warning: false
#| code-fold: true

ie_combined %>%
  filter(Year == latest_year, !is.na(City), !is.na(Rank)) %>%
  mutate(City = str_to_title(City)) %>%
  # Sadece program sayısı fazla olan büyük şehirleri alıyoruz
  group_by(City) %>%
  filter(n() >= 3) %>%
  ungroup() %>%
  # Median Rank'a göre sıralıyoruz (Düşük rakam en iyisi)
  ggplot(aes(x = reorder(City, -Rank, FUN = median, na.rm = TRUE), y = Rank, fill = City)) +
  geom_boxplot(alpha = 0.8, show.legend = FALSE, color = "#94a3b8") +
  scale_fill_viridis_d(option = "mako") +
  scale_y_reverse() +
  coord_flip() + # İsimler daha rahat okunsun diye yatay çevirdik
  theme_minimal() +
  labs(
    title = "Regional Drivers: City Influence on Rankings",
    subtitle = paste("Base Rank Distribution by City in", latest_year),
    x = "City",
    y = "National Base Rank (Lower is Better)"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 16, color = "#38bdf8"),
    plot.subtitle = element_text(size = 12, color = "#cbd5e1", margin = margin(b = 15)),
    axis.text = element_text(color = "#f8fafc", size = 11),
    axis.title = element_text(face = "bold", color = "#f8fafc", size = 12),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_line(color = "#334155", linetype = "dotted")
  )
```

<br>

<div class="glass-card">
## 6. Advanced Correlation: Popularity Drivers vs. Success
To truly decode the drivers of success, we establish a multi-variable correlation model. We analyze how **Demand Ratio** (Total Preferences / Quota) and **Top-Choice Exclusivity** (Students who put the program as their #1 choice) correlate with the final National Rank.
</div>

```{r}
#| label: advanced-correlation
#| message: false
#| warning: false
#| code-fold: true

ie_combined %>%
  filter(Year == latest_year, !is.na(Rank), !is.na(Demand_Ratio)) %>%
  ggplot(aes(x = Demand_Ratio, y = Rank, color = University_Type)) +
  geom_point(aes(size = Top1_Pref), alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed", color = "#f8fafc") +
  scale_color_manual(values = c("State" = "#38bdf8", "Foundation" = "#f472b6")) +
  scale_y_reverse() + 
  theme_minimal() +
  labs(
    title = "Correlation: Demand Intensity vs. Academic Success",
    subtitle = "Higher demand per quota strongly correlates with better (lower) national ranks.",
    x = "Demand Ratio (Total Preferences / Quota)",
    y = "National Base Rank (Lower is Better)",
    size = "Top 1 Choices",
    color = "University Type",
    caption = "Data: Kaggle (Turkish University Admissions)"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 16, color = "#38bdf8"),
    plot.subtitle = element_text(size = 12, color = "#cbd5e1", margin = margin(b = 15)),
    axis.text = element_text(color = "#f8fafc", size = 11),
    axis.title = element_text(face = "bold", color = "#f8fafc", size = 12),
    legend.title = element_text(color = "#f8fafc", face = "bold"),
    legend.text = element_text(color = "#cbd5e1"),
    legend.position = "right",
    panel.grid.major = element_line(color = "#334155", linetype = "dotted"),
    panel.grid.minor = element_blank()
  )
```

---

## 7. Phase 1 Conclusion
As outlined in our project methodology, we successfully harmonized multi-year data and built early analytical models proving the immense impact of geographical location and demand density on national rankings. In Phase 2, we aim to augment this robust baseline by extracting and integrating faculty size metrics and international mobility (Erasmus) data to finalize our comprehensive correlation model.