---
title: "Data Analytics & Results"
subtitle: "Evidence-Based Insights into Industrial Engineering Programs"
sidebar: false
---

<style>
/* DT Tablosunu Karanlık Temaya (Slate) Uyduran Sihirli CSS */
.dataTables_wrapper .dataTables_length, 
.dataTables_wrapper .dataTables_filter, 
.dataTables_wrapper .dataTables_info, 
.dataTables_wrapper .dataTables_paginate {
  color: #cbd5e1 !important;
}
.dataTables_wrapper .dataTables_filter input {
  background-color: #1e293b !important;
  color: #f8fafc !important;
  border: 1px solid #38bdf8 !important;
}
table.dataTable thead th {
  color: #38bdf8 !important; /* Başlıklar Neon Mavi */
  border-bottom: 2px solid #38bdf8 !important;
}
table.dataTable tbody tr {
  background-color: transparent !important;
  color: #f8fafc !important; /* Tablo yazıları beyaz/açık gri */
}
table.dataTable tbody tr:hover {
  background-color: rgba(56, 189, 248, 0.1) !important; /* Üzerine gelince hafif mavi parlama */
}
</style>

### Key Takeaways

* **Dynamic Data Sourcing:** Multi-year dataset (time-series) for Industrial Engineering programs successfully extracted.
* **National Trends:** Clear shifts identified in placement capacities across Turkish universities over the analyzed years.
* **Sector Comparison:** Analysis reveals the structural balance between State and Foundation universities.
* **Strategic Value:** Our interactive dataset allows prospective students and researchers to filter programs by year, type, and institution.

---

<div class="glass-card">
## 1. Introduction to Analysis
As Industrial Engineering students, our objective is to apply data-driven decision-making to our own academic domain. In this section, we transition from raw admission data to actionable insights using `tidyverse` for data wrangling and `ggplot2` for high-impact visualizations.
</div>

## 2. Preliminary Data Exploration & Cleaning

To ensure full reproducibility, we establish a direct query to the database. The raw data contains longitudinal records and distinct program types.

```{r}
#| label: data-processing
#| message: false
#| warning: false
#| code-fold: true

library(tidyverse)
library(DT) 

# 1. Kaggle Verisini Okuma
raw_data <- read_csv("data/01_university_admissions_turkey_2019_2024.csv")

# 2. Veri Temizleme
ie_cleaned <- raw_data %>%
  mutate(
    dept_kucuk = str_to_lower(department_name),
    tur_kucuk = str_to_lower(university_type)
  ) %>%
  filter(str_detect(dept_kucuk, "end") & str_detect(dept_kucuk, "str")) %>%
  mutate(University_Type = case_when(
    str_detect(tur_kucuk, "devlet") ~ "State",
    str_detect(tur_kucuk, "state") ~ "State",
    str_detect(tur_kucuk, "vak") ~ "Foundation",
    str_detect(tur_kucuk, "foundation") ~ "Foundation",
    TRUE ~ "Other"
  )) %>%
  select(year, University_Type, university_name, faculty_name, department_name) %>%
  mutate(university_name = str_squish(university_name)) %>%
  filter(University_Type != "Other")

# 3. İnteraktif Tablo Oluşturma
datatable(
  ie_cleaned, 
  options = list(pageLength = 5, scrollX = TRUE),
  caption = "Table 1: Comprehensive List of Industrial Engineering Programs (2019-2024)"
)
```

<br>

<div class="glass-card">
## 3. Exploratory Data Analysis (EDA)
With a clean dataset, we can now visualize the structural distribution of Industrial Engineering programs. To avoid duplicate counting across multiple years, the plot below isolates the data strictly for the **most recent year** available in the dataset.
</div>

```{r}
#| label: eda-visualizations
#| message: false
#| warning: false
#| code-fold: true

latest_year <- max(as.numeric(ie_cleaned$year), na.rm = TRUE)

ie_cleaned %>%
  filter(year == latest_year) %>%
  count(University_Type) %>%
  ggplot(aes(x = University_Type, y = n, fill = University_Type)) +
  geom_bar(stat = "identity", width = 0.5, show.legend = FALSE, alpha = 0.9) +
  scale_fill_manual(values = c("State" = "#38bdf8", "Foundation" = "#94a3b8")) +
  geom_text(aes(label = n), vjust = -0.5, fontface = "bold", color = "white", size = 5) +
  theme_minimal() +
  labs(
    title = paste("Distribution of IE Programs in", latest_year),
    subtitle = "Active Programs in State vs. Foundation Universities",
    x = "University Sector",
    y = "Number of Active Programs",
    caption = "Data: Kaggle (Turkish University Admissions 2019-2024)"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 16, color = "#38bdf8"),
    plot.subtitle = element_text(size = 12, color = "#cbd5e1", margin = margin(b = 15)),
    axis.text = element_text(color = "#f8fafc", size = 11),
    axis.title = element_text(face = "bold", color = "#f8fafc", size = 12),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(color = "gray", linetype = "dotted"),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA)
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

---

## 4. Conclusion
By isolating the latest academic year, we gain an accurate representation of the current supply of Industrial Engineering departments. Moving forward, we will integrate additional variables such as base scores and regional clusters.