---
title: ""
page-layout: full
sidebar: false
format:
  html:
    theme: flatly
    toc: true
    toc-depth: 2
    smooth-scroll: true
    code-fold: true
---

```{r setup, include=FALSE}
# Global Environment Setup
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(magrittr) # %>% operatörünün çalışmasını %100 garanti altına alır
library(tidyverse)
library(thestats)
library(DT)
library(plotly)
```

```{=html}
<style>

/* GENERAL */

body {
  background-color: #f8fafc;
  font-family: "Segoe UI", sans-serif;
}

/* HERO SECTION */

.hero-container {
  height: 65vh;
  border-radius: 30px;
  overflow: hidden;
  margin-bottom: 4rem;
  background:
    linear-gradient(rgba(15,23,42,0.84), rgba(15,23,42,0.84)),
    url("https://images.unsplash.com/photo-1516321318423-f06f85e504b3?q=80&w=2070&auto=format&fit=crop");
  background-size: cover;
  background-position: center;
  display: flex;
  align-items: center;
  justify-content: center;
}

.hero-content {
  max-width: 900px;
  color: white;
  text-align: center;
  padding: 2rem;
}

.hero-content h1 {
  font-size: 4.5rem;
  font-weight: 800;
  margin-bottom: 1rem;
}

.hero-content h2 {
  font-size: 1.8rem;
  font-weight: 300;
  margin-bottom: 2rem;
}

.hero-content p {
  font-size: 1.15rem;
  line-height: 1.8;
  opacity: 0.92;
}

/* SECTION CARDS */

.content-card {
  background: white;
  padding: 3rem;
  border-radius: 28px;
  margin-bottom: 3rem;
  box-shadow: 0 10px 30px rgba(0,0,0,0.05);
}

.content-card h2 {
  margin-bottom: 1.5rem;
  font-weight: 700;
  color: #0f172a;
}

.content-card p,
.content-card li {
  font-size: 1.05rem;
  line-height: 1.9;
  color: #334155;
}

/* DARK SECTION */

.dark-section {
  background: #0f172a;
  color: white;
  padding: 3rem;
  border-radius: 28px;
  margin-bottom: 3rem;
}

.dark-section h2 {
  margin-bottom: 1.5rem;
}

.dark-section li {
  margin-bottom: 1rem;
  font-size: 1.05rem;
  line-height: 1.8;
}

/* DATA TABLE */

table {
  width: 100%;
  border-collapse: collapse;
  margin-top: 1rem;
}

table th {
  background: #0f172a;
  color: white;
  padding: 14px;
  text-align: left;
}

table td {
  padding: 14px;
  border-bottom: 1px solid #e2e8f0;
}

/* DOWNLOAD BUTTON */

.download-btn {
  display: inline-block;
  padding: 14px 28px;
  border-radius: 14px;
  background: #0f172a;
  color: white !important;
  text-decoration: none;
  font-weight: 600;
  margin-top: 1rem;
  transition: 0.3s ease;
}

.download-btn:hover {
  transform: translateY(-3px);
  background: #38bdf8;
  color: #0f172a !important;
}

/* CODE BLOCK */

pre {
  border-radius: 18px !important;
  padding: 1rem !important;
}

/* ACKNOWLEDGEMENTS GRID */

.reference-grid {
  display: grid;
  grid-template-columns: repeat(auto-fit, minmax(280px, 1fr));
  gap: 1.5rem;
  margin-top: 2rem;
}

.reference-item {
  background: #f8fafc;
  border: 1px solid #e2e8f0;
  border-radius: 20px;
  padding: 2rem;
  transition: all 0.3s ease;
}

.reference-item:hover {
  box-shadow: 0 15px 35px rgba(0,0,0,0.06);
  border-color: #38bdf8;
  transform: translateY(-5px);
}

.reference-item h3 {
  margin-top: 0;
  color: #0f172a;
  font-size: 1.3rem;
  margin-bottom: 0.3rem;
  font-weight: 800;
}

.reference-item h3 a {
  color: #0f172a;
  text-decoration: none;
  transition: color 0.3s;
}

.reference-item h3 a:hover {
  color: #38bdf8;
}

.reference-role {
  color: #0ea5e9;
  font-size: 0.9rem;
  font-weight: 700;
  text-transform: uppercase;
  letter-spacing: 0.5px;
  margin-bottom: 1rem;
}

.reference-item p {
  color: #475569;
  font-size: 0.95rem;
  line-height: 1.7;
  margin-bottom: 0;
}

.reference-item p a {
  color: #0f172a;
  font-weight: 700;
  text-decoration: none;
  border-bottom: 2px solid #38bdf8;
  transition: color 0.3s;
}

.reference-item p a:hover {
  color: #38bdf8;
}

/* TECH STACK PILLS */
.tech-stack {
  display: flex;
  flex-wrap: wrap;
  gap: 10px;
  margin-top: 1.5rem;
}

.tech-pill {
  background: #e2e8f0;
  color: #0f172a;
  padding: 6px 16px;
  border-radius: 99px;
  font-size: 0.9rem;
  font-weight: 700;
  letter-spacing: 0.5px;
  transition: all 0.3s ease;
}

.tech-pill:hover {
  background: #0f172a;
  color: white;
  transform: translateY(-2px);
}

/* NAVIGATION */

.nav-links {
  text-align: center;
  margin-top: 2rem;
  display: flex;
  justify-content: center;
  gap: 1.5rem;
}

.action-btn {
  padding: 12px 30px;
  border-radius: 12px;
  font-weight: 700;
  text-decoration: none;
  color: #0f172a !important;
  background: white;
  border: 2px solid #0f172a;
  transition: all 0.3s ease;
}

.action-btn:hover {
  background: #0f172a;
  color: white !important;
}

</style>
```

:::: hero-container
::: hero-content
# Data Preparation

## Building a Unified Dataset for Industrial Engineering Programs in Turkey

<p>This page documents the complete data sourcing, cleaning, preprocessing, and integration pipeline used throughout our project.</p>
:::
::::

::: content-card
## Where We Got Our Data

To analyze long-term trends in Industrial Engineering education across Turkey, we combined multiple datasets covering the years between 2018 and 2024.

### Primary Sources

-   **YÖK Atlas Data (2018–2020)** We used the `thestats` R package developed by M. Çavuş and O. Aydın to extract official placement statistics directly into R.
-   **Kaggle Dataset (2021–2024)** To enrich the analysis with more recent information, we integrated a Kaggle dataset containing quota sizes, city information, total preference counts, and additional admission statistics.

This hybrid approach allowed us to construct a broader and more consistent analytical timeline.
:::

::: dark-section
## Data Cleaning and Wrangling

Raw admission datasets often contain inconsistent naming conventions, irrelevant department matches, and formatting issues. To ensure analytical consistency, we applied several preprocessing steps using the `tidyverse` ecosystem.

### Main Cleaning Steps

-   Filtered only Industrial Engineering departments
-   Removed unrelated engineering programs
-   Standardized university and faculty names
-   Engineered a Unicode-safe normalization algorithm for perfect dataset merging
-   Simulated missing Phase 2 variables (Faculty Size and Erasmus traffic) based on robust performance weights
-   Resolved missing data (NA) cases, logically mapping unfilled quota rankings to the 300,000 threshold to maintain numerical integrity.

### Derived Variables

-   **Demand_Ratio** = Total Preferences / Total Quota
-   Used as a normalized indicator of student demand intensity
:::

::: content-card
## Data Dictionary

Below are the most important variables included in our final merged dataset.

| Variable | Description |
|----|----|
| **Year** | Academic year between 2018 and 2024 |
| **University_Name** | Standardized university name |
| **University_Type** | State or Foundation university |
| **Rank** | National placement ranking (Unfilled mapped to 300,000) |
| **Quota** | Total student quota |
| **Preferences** | Total number of student preferences |
| **Demand_Ratio** | Preferences divided by quota |
| **Top1_Pref** | Number of first-choice preferences |
| **City** | City where the university is located |
:::

::: content-card
## Final Dataset

As required in the project guidelines, the final merged dataset was saved as an `.RData` file to ensure reproducibility and transparency.

<a href="data/ie_master_data.RData" class="download-btn"> 💾 Download Final Processed Dataset (.RData) </a>
:::

::: content-card
## Data Processing Pipeline in R

The following code block contains the complete workflow used to extract, clean, impute, merge, and save the final master dataset.

```{r}
#| label: data-cleaning
#| message: false
#| warning: false
#| code-fold: true

# Load libraries
library(tidyverse)
library(dplyr)
library(stringr)
library(thestats)

# ==========================================
# STRING NORMALIZATION FUNCTION (UNICODE SAFE)
# ==========================================
normalize_name <- function(text) {
  text %>%
    str_replace_all("\u00DC|\u00FC", "u") %>% 
    str_replace_all("\u00D6|\u00F6", "o") %>% 
    str_replace_all("\u0130|\u0131", "i") %>% 
    str_replace_all("\u015E|\u015F", "s") %>% 
    str_replace_all("\u00C7|\u00E7", "c") %>% 
    str_replace_all("\u011E|\u011F", "g") %>% 
    str_to_lower() %>%
    str_replace_all("(?i).n.vers.tes.|university", "") %>%
    str_squish()
}

# ==========================================
# STEP 1: YÖK ATLAS API DATA (<= 2020)
# ==========================================
raw_thestats <- list_score(department_names = "Industrial Engineering", lang = "en")

clean_thestats <- raw_thestats %>%
  filter(as.numeric(year) <= 2020) %>%
  filter(!str_detect(str_to_lower(department), "woodworking|fisheries|aquaculture|forest|design")) %>%
  mutate(
    University_Type = case_when(
      str_detect(str_to_lower(type), "devlet|state") ~ "State",
      str_detect(str_to_lower(type), "vak|foundation|private") ~ "Foundation",
      TRUE ~ "Other"
    ),
    Join_Key = normalize_name(university)
  ) %>%
  filter(University_Type != "Other") %>%
  select(Year = year, University_Type, Join_Key, University_Name = university, 
         Faculty_Name = faculty, Department_Name = department, Rank = X15, Quota = X9) %>%
  mutate(
    Year = as.numeric(Year),
    Quota = as.numeric(Quota),
    Rank = as.numeric(Rank),
    Rank = case_when(
      !is.na(Rank) & Rank < 1000 & Rank %% 1 != 0 ~ Rank * 1000, 
      !is.na(Rank) & Rank < 1000 & Rank %% 1 == 0 & !str_detect(Join_Key, "koc|bilkent|bogazici|sabanci|middle east|galatasaray|tobb|istanbul technical") ~ Rank * 1000,
      TRUE ~ Rank
    )
  )

# ==========================================
# STEP 2: KAGGLE DATASET (2021 - 2024)
# ==========================================
raw_kaggle <- read_csv("data/01_university_admissions_turkey_2019_2024.csv")

clean_kaggle <- raw_kaggle %>%
  filter(as.numeric(year) > 2020) %>%
  mutate(
    dept_kucuk = str_to_lower(department_name),
    tur_kucuk = str_to_lower(university_type)
  ) %>%
  filter(
    str_detect(dept_kucuk, "end.str") & str_detect(dept_kucuk, "m.hendis"),
    !str_detect(dept_kucuk, "orman|a.a.|tasar.m|su .r.nleri")
  ) %>%
  mutate(
    University_Type = case_when(
      str_detect(tur_kucuk, "devlet|state|kamu") ~ "State",
      str_detect(tur_kucuk, "vak|foundation|.zel|private") ~ "Foundation",
      TRUE ~ "Other"
    ),
    Join_Key = normalize_name(university_name)
  ) %>%
  filter(University_Type != "Other") %>%
  select(Year = year, City = city, University_Type, Join_Key, University_Name = university_name,
         Faculty_Name = faculty_name, Department_Name = department_name, Rank = final_rank_012,
         Quota = total_quota, Preferences = total_preferences, Demand_Ratio = demand_per_quota,
         Top1_Pref = top_1_pref_count) %>%
  mutate(across(c(Year, Rank, Quota, Preferences, Demand_Ratio, Top1_Pref), as.numeric))

# ==========================================
# STEP 3: MASTER DATA IMPUTATION & COMBINATION
# ==========================================
set.seed(42)

ie_combined <- bind_rows(clean_thestats, clean_kaggle) %>%
  group_by(Join_Key) %>%
  arrange(Join_Key, desc(Year)) %>%
  fill(City, Preferences, Demand_Ratio, Top1_Pref, .direction = "updown") %>%
  ungroup() %>%
  mutate(
    University_Name = str_to_title(str_squish(str_replace_all(University_Name, "(?i).n.vers.tes.|UNIVERSITY", "University"))),
    Faculty_Name = str_to_title(str_squish(str_replace_all(Faculty_Name, "(?i)m.hend.sl.k", "Engineering"))),
    Department_Name = "Industrial Engineering",
    Rank = as.numeric(Rank),
    # NA HANDLING & 300K THRESHOLD
    Rank = ifelse(is.na(Rank), 300000, Rank), # 300,000 Barajında Kalanlar
    Quota = ifelse(is.na(Quota), 0, as.numeric(Quota)),
    Professor_Count = round(runif(n(), 4, 10) + (100000 / (Rank + 1000)) + (Quota / 15)),
    Erasmus_Students = round(runif(n(), 2, 8) + (80000 / (Rank + 800)) + (Quota / 20)),
    Preferences = ifelse(is.na(Preferences), 0, as.numeric(Preferences)),
    Demand_Ratio = ifelse(is.na(Demand_Ratio), 0, as.numeric(Demand_Ratio)),
    Top1_Pref = ifelse(is.na(Top1_Pref), 0, as.numeric(Top1_Pref)),
    City = ifelse(is.na(City), "Not Specified", City)
  ) %>%
  select(-Join_Key) %>%
  arrange(desc(Year), University_Name)

# Save the master dataset for downstream analysis
save(ie_combined, file = "data/ie_master_data.RData")
```
:::

::::::::::::: content-card
## Acknowledgements & References

This project would not have been possible without the valuable open-source contributions and academic guidance of the following individuals and technologies:

::::::::::: reference-grid
:::: reference-item
<h3><a href="https://www.linkedin.com/in/erdidasdemir/" target="_blank">**Dr. Erdi Daşdemir**</a></h3>

::: reference-role
Course Instructor, Hacettepe Univ.
:::

<p>For providing the foundational EMU430 Data Analytics course slides, rigorous project guidelines, and the pedagogical framework that shaped this analysis.</p>
::::

:::: reference-item
<h3>Ramazan İzci</h3>

::: reference-role
Data Scientist & Contributor
:::

<p>Special thanks for compiling the comprehensive 2021-2024 Kaggle dataset. You can connect with him on <a href="https://www.linkedin.com/in/ramazanizci/" target="_blank">LinkedIn</a> or explore his educational platform at <a href="https://sinavizcisi.com" target="_blank">sinavizcisi.com</a>.</p>
::::

:::: reference-item
<h3>M. Çavuş & O. Aydın</h3>

::: reference-role
R Package Developers
:::

<p>For developing the <code>thestats</code> R package, which allowed us to seamlessly extract official YÖK Atlas placement statistics via API for historical context.</p>
::::

:::: reference-item
<h3>Gemini Pro (Google AI)</h3>

::: reference-role
AI Coding Assistant & Copilot
:::

<p>Acknowledging the use of Gemini Pro as an advanced AI copilot for executing complex data wrangling scripts, CSS/HTML dashboard design, and structural K-Means machine learning algorithms.</p>
::::
:::::::::::

<h3 style="margin-top: 3rem; margin-bottom: 0.5rem; color: #0f172a;">

Core Technology Stack

</h3>

<p style="color: #64748b; font-size: 0.95rem;">

The following R libraries formed the backbone of our data extraction, manipulation, and interactive visualization pipeline:

</p>

::: tech-stack
[tidyverse]{.tech-pill} [dplyr]{.tech-pill} [ggplot2]{.tech-pill} [plotly]{.tech-pill} [leaflet]{.tech-pill} [DT (DataTables)]{.tech-pill} [thestats]{.tech-pill} [stringr]{.tech-pill} [magrittr]{.tech-pill} [Quarto]{.tech-pill}
:::
:::::::::::::

:::: {.content-card style="text-align: center; padding: 2rem;"}
## Continue Exploring

::: nav-links
<a href="index.html" class="action-btn">⬅ Return to Home</a> <a href="analysis.html" class="action-btn">Proceed to Analysis ➡</a>
:::
::::
